---
title: 'RR Peer Assessment #1'
author: "Suzanne"
date: "Sunday, April 12, 2015"
output: html_document
---
##Peer Assessment #1

Let's start by loading and processing the data:
```{r}
setInternet2(TRUE)
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", temp, mode="wb")
ad <- read.csv(unz(temp,"activity.csv"),stringsAsFactors=FALSE)
ad$date <- as.Date(ad$date)
```

Here is a dataframe showing the total steps taken by day:
```{r, echo=TRUE}
dailySteps <- aggregate(steps~date, data=ad,FUN=sum, na.rm=TRUE)
head(dailySteps)
```

Here is a histogram of total steps taken in a day:
```{r, echo = TRUE}
hist(dailySteps$steps)
```

Here is the mean number of steps per day
```{r}
dailyMean <- round(mean(dailySteps$steps),0)
dailyMean
```
Here is the median number steps per day
```{r}
dailyMedian <- median(dailySteps$steps)
dailyMedian
```


Here is a time series of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
intervalSteps <- aggregate(steps~interval, data=ad,FUN=mean, na.rm=TRUE)
plot(intervalSteps$interval, intervalSteps$steps, type="l")
```

```{r}
#interval with largest avg steps
maxInterval<-intervalSteps[which.max(intervalSteps$steps),1]
```

The interval with the largest average steps is `r maxInterval`.

```{r}
#number of complete rows
missingValues <- nrow(na.omit(ad))
sum(complete.cases(ad))
```

There are `r missingValues` missing values in this dataset.

Let's create new dataset by imputting missing values using the average of that interval
```{r}
#merge the original dataset with the dataframe containing average steps per interval
combo <- merge(ad, intervalSteps, by="interval")
colnames(combo)[4] <- "intervalAvg"

#replace NULLs with the interval avg
combo$steps.x[is.na(combo$steps.x)] <- combo$intervalAvg[is.na(combo$steps.x)]

#remove the interval avg column after imputting
combo$intervalAvg <-  NULL
```

Let's make histogram of total steps taken per day from the imputted dataset:
```{r}
comboDailySteps <- aggregate(steps.x~date, data=combo,FUN=sum, na.rm=TRUE)

#calculate mean and median steps per day
comboDailyMean <- mean(comboDailySteps$steps.x)
comboDailyMedian <- median(comboDailySteps$steps.x)
comboDailyMean
comboDailyMedian
```

Comparing the histograms of the two datasets shows there is an impact of imputting missing values, by raising the frequency of occurences in the 10,000 to 15,000 step range:
```{r}
par(mfrow=c(1,2))
hist(comboDailySteps$steps.x, main = "Histogram With Imputted Values")
hist(dailySteps$steps, main = "Histogram with Missing Values")
```


Create new variable indicating weekday vs. weekend
```{r}
combo$DayofWeek <- ifelse(weekdays(combo$date) %in% c('Saturday','Sunday'), "weekend", "weekday")
```



Aggregate steps by interval for weekday. Aggregate steps by interval for weekends. Compare the results
```{r}
comboIntervalSteps <- aggregate(steps.x~interval + DayofWeek, data=combo,FUN=sum, na.rm=TRUE)
par(mfrow=c(1,2))
with(comboIntervalSteps[comboIntervalSteps$DayofWeek == "weekend",], plot(interval,steps.x, type="l", main=" Weekend Intervals"))
with(comboIntervalSteps[comboIntervalSteps$DayofWeek == "weekday",], plot(interval,steps.x, type="l", main = "Weekday Interval"))
```

